### 2022-04-07 chapters/fairness/fvgm/abstract.tex
updated files

\section{abstract}
In recent years, machine learning (ML) algorithms have been deployed in safety-critical and high-stake decision-making, where the \textit{fairness} of algorithms is of paramount importance. Fairness in ML  centers on detecting bias towards certain demographic populations induced by an ML classifier and proposes algorithmic solutions to mitigate the bias with respect to different fairness definitions.  To this end, several \textit{fairness verifiers} have been proposed that compute the bias in the prediction of an ML classifier\textemdash essentially beyond a finite dataset\textemdash given the probability distribution of input features. In the context of verifying \textit{linear classifiers}, existing fairness verifiers are limited by \emph{accuracy} due to imprecise modeling of correlations among features and \emph{scalability} due to restrictive formulations of the classifiers as SSAT/SMT formulas or by sampling.

In this paper, we propose an efficient fairness verifier, called \fvgm, that encodes the correlations among features as a Bayesian network. In contrast to existing verifiers, \fvgm~proposes a \textit{stochastic subset-sum} based approach for verifying linear classifiers. Experimentally, we show that  \fvgm~leads to an \textit{accurate} and \textit{scalable} assessment for more diverse families of fairness-enhancing algorithms, fairness attacks, and group/causal fairness metrics than the state-of-the-art. We also demonstrate that {\fvgm} facilitates the computation of fairness influence functions as a stepping stone to detect the source of bias induced by subsets of features.

>>>>

\begin{abstract}
In recent years, machine learning (ML) algorithms have been deployed in safety-critical and high-stake decision-making, where the \textit{fairness} of algorithms is of paramount importance. Fairness in ML  centers on detecting bias towards certain demographic populations induced by an ML classifier and proposes algorithmic solutions to mitigate the bias with respect to different fairness definitions.  To this end, several \textit{fairness verifiers} have been proposed that compute the bias in the prediction of an ML classifier\textemdash essentially beyond a finite dataset\textemdash given the probability distribution of input features. In the context of verifying \textit{linear classifiers}, existing fairness verifiers are limited by \emph{accuracy} due to imprecise modeling of correlations among features and \emph{scalability} due to restrictive formulations of the classifiers as SSAT/SMT formulas or by sampling.

In this paper, we propose an efficient fairness verifier, called \fvgm, that encodes the correlations among features as a Bayesian network. In contrast to existing verifiers, \fvgm~proposes a \textit{stochastic subset-sum} based approach for verifying linear classifiers. Experimentally, we show that  \fvgm~leads to an \textit{accurate} and \textit{scalable} assessment for more diverse families of fairness-enhancing algorithms, fairness attacks, and group/causal fairness metrics than the state-of-the-art fairness verifiers. We also demonstrate that {\fvgm} facilitates the computation of fairness influence functions as a stepping stone to detect the source of bias induced by subsets of features.
\end{abstract}

### 2022-04-07 chapters/fairness/fvgm/applications.tex
updated files

\section{Applications of {\fvgm}}\label{fairness_fvgm_sec:applications}
In this section, we apply {\fvgm} for verifying fairness-enhancing algorithms and depreciation attacks. We also show that {\fvgm} facilitates computation of fairness influence functions enabling detection of bias due to individual features.

@@

	\centering
	\subfloat{\includegraphics[scale=0.25]{figures/fairness/fvgm/disp_fairness_attack}}
	\subfloat{\includegraphics[scale=0.25]{figures/fairness/fvgm/stat_fairness_attack}}
	%\vspace{-0.5em}
	\caption{Verifying poisoning attack against fairness using {\fvgm}. The red line denotes the safety margin of the ML model against the attack.}\label{fairness_fvgm_fig:fairness_attacks}
\end{figure}
\fi

\paragraph{Verifying Fairness-enhancing Algorithms.} We deploy {\fvgm} in verifying the effectiveness of fairness-enhancing algorithms designed to ameliorate bias. For example, fairness 
pre-processing algorithms can be validated by applying {\fvgm} on the unprocessed and the processed data separately and comparing different fairness metrics. In Table~\ref{fairness_fvgm_tab:fair_algo_verification}, we report the effect of fairness algorithms w.r.t. four fairness metrics: disparate impact (DI), path-specific causal fairness (PCF), statistical parity (SP), and equalized odds (EO). Note that, fairness is improved if DI increases and the rest of the metrics decrease. In most instances for Adult dataset, reweighing (RW)~\cite{kamiran2012data} and optimized pre-processing (OP)~\cite{calmon2017optimized} algorithms are successful in improving fairness. Except for the unfairness regarding the sensitive feature `sex', where OP algorithm fails. Thus, \textit{{\fvgm} verifies the enhancement and decrement in fairness by  fairness-enhancing algorithms. }

\paragraph{Verifying Fairness Attacks.}
We apply {\fvgm} in verifying a fairness poisoning-attack algorithm. This algorithm injects a small fraction of poisoned samples into the training data such that the classifier becomes relatively unfair~\cite{solans2020poisoning}. We apply this attack to add $\{1, 5, \dots, 160\}$ poisoned samples  and measure the corresponding DI's. In Figure~\ref{fairness_fvgm_fig:fairness_attacks},  {\fvgm} verifies that DI of the classifier decreases, i.e.\textit{ the classifier becomes more unfair},\textit{ as the number of poisoned samples increases}. Therefore, {\fvgm} shows the potential of being deployed in safety-critical applications to detect fairness attacks. For example, if we set $0.9$ as an acceptable threshold of DI, {\fvgm} can raise an alarm once $160$ poisoned samples are added.

@@

	\centering
	\small
	\caption{Verification of fairness algorithms using {\fvgm}. Numbers in bold refer to fairness improvement. RW and OP refer to reweighing and optimized-preprocessing algorithms.  }\label{fairness_fvgm_tab:fair_algo_verification}
	\setlength{\tabcolsep}{.4em}

@@

\paragraph{Fairness Influence Function (FIF): Tracing Sources of Unfairness.}
Another application of \fvgm{} as a fairness verifier is to quantify the effect of a subset of features on fairness. Thus, we define fairness influence function (FIF)  that computes the effect of a subset of features $\mathbf{S}$ on the PPV of a classifier given a specific sensitive group $ \sensitive = \mathbf{a}$,
$
	\mathsf{FIF}(\mathbf{S}) \triangleq \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}, \mathcal{D}] - \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a},  \mathcal{D}_{-\mathbf{S}}].
$
FIF allows us to explain the sources of unfairness in the classifier. In practice, we compute FIF of $\mathbf{S}$ by replacing the probability distribution of $\mathbf{S}$ with a uniformly random distribution, referred to as  $ \mathcal{D}_{-\mathbf{S}} $, and reporting differences in the conditional PPVs. Further details of FIF are in Appendix.

In Figure~\ref{fairness_fvgm_fig:influence function}, we compute FIF for all features in COMPAS dataset, separately for two sensitive groups: Female (`sex' $ = 1 $) and Male. This dataset concerns the likelihood of a person of re-offending crimes within the next two years. We first observe that the base PPV values are different for the two groups ($ 0.46 $ vs $ 0.61 $ for Female and Male), thereby showing Male as more probable to re-offend crimes than Female. Moreover, FIF of the feature `age' is comparatively higher in magnitude for Male than Female. 
This implies that while deciding recidivism the algorithm assumes that Female individuals across different ages re-offend crimes with almost the same probability and the probability of re-offending for Male individuals highly depends on age. Thus, applying {\fvgm} to compute FIF provides us insights about sources of bias and thus, indicators to improve fairness.

\begin{figure}[t!]	
	\centering
	\vspace{-2ex}
	\subfloat{\includegraphics[scale=0.23]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_1}}
	\subfloat{\includegraphics[scale=0.23]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_0}}
	%\vspace{-0.5em}
%	\vspace{-2ex}
	\caption{\footnotesize FIF computation for COMPAS dataset. For Female (left plot) and Male, influence of `age' decreases and the PPV of the classifier increases  by different magnitudes.}\label{fairness_fvgm_fig:influence function}
%	\vspace{-2ex}
\end{figure}

>>>>

\section{Applications of {\fvgm}}\label{fvgm_sec:applications}
In this section, we apply {\fvgm} for verifying fairness-enhancing algorithms and depreciation attacks. We also demonstrate that {\fvgm} facilitates computation of fairness influence functions by enabling the detection of bias due to individual features.

@@

	\centering
	\subfloat{\includegraphics[scale=0.25]{figures/fairness/fvgm/disp_fairness_attack}}
	\subfloat{\includegraphics[scale=0.25]{figures/fairness/fvgm/stat_fairness_attack}}
	%\vspace{-0.5em}
	\caption{Verifying poisoning attack against fairness using {\fvgm}. The red line denotes the safety margin of the ML model against the attack.}\label{fvgm_fig:fairness_attacks}
\end{figure}
\fi

\paragraph{Verifying Fairness-enhancing Algorithms.} We deploy {\fvgm} in verifying the effectiveness of fairness-enhancing algorithms designed to ameliorate bias. For example, fairness 
pre-processing algorithms can be validated by applying {\fvgm} on the unprocessed and the processed data separately and comparing different fairness metrics. In Table~\ref{fvgm_tab:fair_algo_verification}, we report the effect of fairness algorithms w.r.t. four fairness metrics: disparate impact (DI), path-specific causal fairness (PCF), statistical parity (SP), and equalized odds (EO). Note that, fairness is improved if DI increases and the rest of the metrics decrease. For instance, in most instances for Adult dataset, reweighing (RW)~\cite{kamiran2012data} and optimized pre-processing (OP)~\cite{calmon2017optimized} algorithms are successful in improving fairness. The exceptional case is the unfairness regarding the sensitive feature `sex', where OP algorithm fails in improving fairness metric EO. Thus, \textit{{\fvgm} verifies the enhancement and decrement in fairness by  fairness-enhancing algorithms. }

\begin{figure}
	\centering
	\subfloat{\includegraphics[scale=0.2]{figures/fairness/fvgm/disp_fairness_attack}}
	\subfloat{\includegraphics[scale=0.2]{figures/fairness/fvgm/stat_fairness_attack}}

	\captionof{figure}{Verifying poisoning attack against fairness using {\fvgm}. The red line denotes the safety margin of the ML model against the attack.}\label{fvgm_fig:fairness_attacks}
\end{figure}

\paragraph{Verifying Fairness Attacks.}
We apply {\fvgm} in verifying a fairness poisoning-attack algorithm. This algorithm injects a small fraction of poisoned samples into the training data such that the classifier becomes relatively unfair~\cite{solans2020poisoning}. We apply this attack to add $\{1, 5, \dots, 160\}$ poisoned samples  and measure the corresponding disparate impact and statistical parity. In Figure~\ref{fvgm_fig:fairness_attacks},  {\fvgm} verifies that the disparate impact of the classifier decreases and statistical parity increases, i.e.\textit{ the classifier becomes more unfair},\textit{ as the number of poisoned samples increases}. Therefore, {\fvgm} shows the potential of being deployed in safety-critical applications to detect fairness attacks. For example, if we set $0.9$ as an acceptable threshold of disparate impact, {\fvgm} can raise an alarm once $160$ poisoned samples are added.

@@

	\centering
	\small
	\caption{Verification of fairness algorithms using {\fvgm}. Numbers in bold refer to fairness improvement. RW and OP refer to reweighing and optimized-preprocessing algorithms.  }\label{fvgm_tab:fair_algo_verification}
	\setlength{\tabcolsep}{.4em}

@@

\paragraph{Fairness Influence Function (FIF): Tracing Sources of Unfairness.}
Another application of \fvgm{} as a fairness verifier is to quantify the effect of a subset of features on fairness. Thus, we define fairness influence function (FIF)  that computes the effect of a subset of non-sensitive features $\mathbf{S} \subseteq \nonsensitive$ on the PPV of a classifier given a specific sensitive group $ \sensitive = \mathbf{a}$,
$
	\mathsf{FIF}(\mathbf{S}) \triangleq \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}, \mathcal{D}] - \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a},  \mathcal{D}_{-\mathbf{S}}].
$
FIF allows us to explain the sources of unfairness in the classifier. In practice, we compute FIF of $\mathbf{S}$ by replacing the probability distribution of $\mathbf{S}$ with a uniformly random distribution, referred to as  $ \mathcal{D}_{-\mathbf{S}} $, and reporting differences in the conditional PPVs.

In Figure~\ref{fvgm_fig:influence function}, we compute FIF for all features in COMPAS dataset, separately for two sensitive groups: Female (`sex' $ = 1 $) and Male. This dataset concerns the likelihood of a person of re-offending crimes within the next two years. We first observe that the base PPV values are different for the two groups ($ 0.46 $ vs $ 0.61 $ for Female and Male), thereby showing Male as more probable to re-offend crimes than Female. Moreover, FIF of the feature `age' is comparatively higher in magnitude for Male than Female. 
This implies that while deciding recidivism the algorithm assumes that Female individuals across different ages re-offend crimes with almost the same probability and the probability of re-offending for Male individuals highly depends on age. Thus, applying {\fvgm} to compute FIF provides us insights about sources of bias and thus, indicators to improve fairness.

\begin{figure}[t!]	
	\centering
	\vspace{-2ex}
	\subfloat{\includegraphics[scale=0.23]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_1}}
	\subfloat{\includegraphics[scale=0.23]{figures/fairness/fvgm/dependency_exp_Learn-dependency_compas_lr_sex_0}}
	%\vspace{-0.5em}
%	\vspace{-2ex}
	\caption{FIF computation for COMPAS dataset. For Female (left plot) and Male, influence of `age' decreases and the PPV of the classifier increases  by different magnitudes.}\label{fvgm_fig:influence function}
%	\vspace{-2ex}
\end{figure}

### 2022-04-07 chapters/fairness/fvgm/background.tex
updated files

 	\begin{enumerate}[leftmargin=*]
 		\itemsep0em 
 		\item \textit{Disparate impact} (DI): A classifier  satisfies $(1 - \epsilon)$-disparate impact if for $\epsilon \in [0,1] $,
 		$
 		\min_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}]  \ge (1 - \epsilon) \max_{\mathbf{a}'} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}'].
 		$
 		\item \textit{Statistical Parity} (SP): A classifier satisfies $\epsilon$-statistical parity if for $\epsilon \in [0,1] $, 
 		$
 		\max_{\mathbf{a}, \mathbf{a}'}|\Pr[\hat{Y} =1 | \mathbf{A} = \mathbf{a}] - \Pr [\hat{Y} = 1| \mathbf{A} = \mathbf{a}']| \le \epsilon.
 		$
 		\item \textit{Equalized Odds} (EO): 	A classifier satisfies $\epsilon$-equalized odds if for $\epsilon \in [0,1] $,
 		$ \max_{\mathbf{a}, \mathbf{a}'} |\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 0  ] - \Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}', Y = 0]| \le \epsilon, $ and $
 		\max_{\mathbf{a}, \mathbf{a}'}|\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 1  ] - \Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}', Y = 1]| \le \epsilon.
 		$
 		\item \textit{Path-specific Causal Fairness} (PCF): 
 		Let $ \mathbf{a}_{\max}  \triangleq \argmax_{ \mathbf{a}} \Pr[\hat{Y} =1 |\mathbf{A}=  \mathbf{a}] $. We consider mediator features $ \mediator \subseteq \nonsensitive $ sampled from the conditional distribution $ {\mathcal{Z}_{|\mathbf{A} = \mathbf{a}_{\max}}} $. This emulates the fact that mediator variables have the same sensitive features $ \mathbf{a}_{\max} $.  For $ \epsilon \in [0,1] $,  path-specific causal fairness is defined as 
 		$
 		\max_{\mathbf{a}, \mathbf{a}'} |\Pr[\hat{Y} = 1 | \sensitive =  \mathbf{a}, \mediator] - \Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}', \mediator ]| \le \epsilon
 		$.
 	\end{enumerate}

 	  For all of the above metrics, lower value of $\epsilon$ indicates higher fairness demonstrated by the classifier $\mathcal{M}$. Following the observation of~\cite{ghosh2020justicia},  computing all of the aforementioned fairness metrics is equivalent to computing the maximum and minimum of the \textit{Positive Predictive Value} (PPV) of the classifier, denoted as $\Pr[\hat{Y}=1|\mathbf{A} =\mathbf{a}]$, for all compound sensitive groups $\mathbf{a}$ from $ \sensitive $. Thus, in Section~\ref{fairness_fvgm_sec:fvgm}, we focus on computing the maximum and minimum PPVs and then extend it to assess corresponding fairness metrics. We call the group for which PPV is maximum (minimum) as the \textit{most (least) favored group}.

@@

	\textbf{Stochastic Subset Sum Problem ({\stochastic}).} 
	Let $ \mathbf{B} \triangleq \{B_i\}_{i=1}^{|\mathbf{B}|}$ be a set of Boolean variables and $ w_i \in \mathbb{Z} $ be the weight of $ B_i $. Given a constraint of the form  $\sum_{i = 1}^ {|\mathbf{B}|} w_i B_i = \tau $, for a constant $ \tau \in \mathbb{Z} $, the subset-sum problem seeks to compute $\mathbf{b} \in \{0,1\}^{|\mathbf{B}|}$ such that the constraint evaluates to true when $\mathbf{B}$ is substituted with $\mathbf{b}$. Subset sum problem is known to be a $ \mathrm{NP} $-complete problem and well-studied in theoretical computer science~\cite{kleinberg2006algorithm}. The \textit{counting} version of the subset-sum problem counts all $ \mathbf{b} $'s for which the above constraint holds; and this problem belongs to the complexity class $ \mathrm{\#P} $. In this paper, we consider the counting problem for the constraint $\sum_{i = 1}^ {|\mathbf{B}|} w_i B_i \ge \tau $ where variables $ B_i $'s are stochastic. More precisely, we define a \textit{stochastic subset-sum} problem, namely {\stochastic}, that computes $ \Pr[\sum_{i = 1}^ {|\mathbf{B}|} w_iB_i \ge \tau] $.    Details of {\stochastic} are in Section~\ref{fairness_fvgm_sec:stochastic_sum_set_sum}.

	\textbf{Bayesian Network.}
	In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In this paper, we leverage an access to a Bayesian network on $(\nonsensitive, \sensitive)$ that represents the joint distribution on them. 
	A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\mathbf{V}, \mathbf{E}) $ is a Directed Acyclic Graph (DAG), and $\theta$ is a set of parameters that encodes the conditional probabilities induced by the joint distribution under investigation. Each vertex $V_i \in \mathbf{V}$ corresponds to a random variable. Edges $ \mathbf{E} \in \mathbf{V} \times \mathbf{V} $ imply conditional dependencies among variables. For each variable $ V_i \in \mathbf{V} $, let $ \parent(V_i) \subseteq \mathbf{V} \setminus \{V_i\} $ denote the set of parents of $ V_i $. Given $\parent(V_i)$ and parameters $\theta$, $ V_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ V_i $ and $ \mathbf{u} $ of $ \parent(V_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\mathbf{V}$ as:

	\begin{align}
	\Pr[V_1=v_1, \dots, V_{|\mathbf{V}|}=v_{|\mathbf{V}|}] = \prod_{i=1}^{|\mathbf{V}|} \Pr[V_i = v_i | \parent(V_i) = \mathbf{u}; \theta].
	\label{fairness_fvgm_eq:BN}
	\end{align}

>>>>

 	\begin{enumerate}[leftmargin=*]
 		\itemsep0em 
 		\item \textit{Disparate Impact} (DI): A classifier  satisfies $(1 - \epsilon)$-disparate impact if for $\epsilon \in [0,1] $,
 		$
 		\min_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}]  \ge (1 - \epsilon) \max_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}].
 		$
 		\item \textit{Statistical Parity} (SP): A classifier satisfies $\epsilon$-statistical parity if for $\epsilon \in [0,1] $, 
 		$
 		\max_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}] - \min_{\mathbf{a}} \Pr[\hat{Y} =1 | \mathbf{A} =  \mathbf{a}] \le \epsilon.
 		$
 		\item \textit{Equalized Odds} (EO): 	A classifier satisfies $\epsilon$-equalized odds if for $\epsilon \in [0,1] $,
 		$ \max_{\mathbf{a}}\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 0  ] - \min_{\mathbf{a}}\Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}, Y = 0] \le \epsilon, $ and $
 		\max_{\mathbf{a}}\Pr[\hat{Y} =1 |\mathbf{A}= \mathbf{a}, Y= 1  ] - \min_{\mathbf{a}}\Pr [\hat{Y} = 1|\mathbf{A}= \mathbf{a}, Y = 1] \le \epsilon.
 		$
 		\item \textit{Path-specific Causal Fairness} (PCF): 
 		Let $ \mathbf{a}_{\max}  \triangleq \argmax_{ \mathbf{a}} \Pr[\hat{Y} =1 |\mathbf{A}=  \mathbf{a}] $. We consider mediator features $ \mediator \subseteq \nonsensitive $ sampled from the conditional distribution $ {\mathcal{Z}_{|\mathbf{A} = \mathbf{a}_{\max}}} $. This emulates the fact that mediator variables have the same sensitive features $ \mathbf{a}_{\max} $.  For $ \epsilon \in [0,1] $,  path-specific causal fairness is defined as 
 		$
 		\max_{\mathbf{a}}\Pr[\hat{Y} = 1 | \sensitive =  \mathbf{a}, \mediator] - \min_{\mathbf{a}}\Pr[\hat{Y} = 1 | \sensitive = \mathbf{a}, \mediator ] \le \epsilon
 		$.
 	\end{enumerate}

 	  For all of the above metrics, lower value of $\epsilon$ indicates higher fairness demonstrated by the classifier $\mathcal{M}$. Following the observation of~\cite{ghosh2020justicia},  computing all of the aforementioned fairness metrics is equivalent to computing the maximum and minimum of the \textit{Positive Predictive Value} (PPV) of the classifier, denoted as $\Pr[\hat{Y}=1|\mathbf{A} =\mathbf{a}]$, for all compound sensitive groups $\mathbf{a}$ from $ \sensitive $. Thus, in Section~\ref{fvgm_sec:fvgm}, we focus on computing the maximum and minimum PPVs and then extend it to assess corresponding fairness metrics. We call the group for which PPV is maximum (minimum) as the \textit{most (least) favored group}.

@@

	\textbf{Stochastic Subset Sum Problem ({\stochastic}).} 
	Let $ \mathbf{B} \triangleq \{B_i\}_{i=1}^{|\mathbf{B}|}$ be a set of Boolean variables and $ w_i \in \mathbb{Z} $ be the weight of $ B_i $. Given a constraint of the form  $\sum_{i = 1}^ {|\mathbf{B}|} w_i B_i = \tau $, for a constant threshold $ \tau \in \mathbb{Z} $, the subset-sum problem seeks to compute an assignment $\mathbf{b} \in \{0,1\}^{|\mathbf{B}|}$ such that the constraint evaluates to true when $\mathbf{B}$ is substituted with $\mathbf{b}$. Subset sum problem is known to be a $ \mathrm{NP} $-complete problem and well-studied in theoretical computer science~\cite{kleinberg2006algorithm}. The \textit{counting} version of the subset-sum problem counts all $ \mathbf{b} $'s for which the above constraint holds; and this problem belongs to the complexity class $ \mathrm{\#P} $. In this paper, we consider the counting problem for the constraint $\sum_{i = 1}^ {|\mathbf{B}|} w_i B_i \ge \tau $ where variables $ B_i $'s are stochastic. More precisely, we define a \textit{stochastic subset-sum} problem, namely {\stochastic}, that computes $ \Pr[\sum_{i = 1}^ {|\mathbf{B}|} w_iB_i \ge \tau] $.    Details of {\stochastic} are in Section~\ref{fvgm_sec:stochastic_sum_set_sum}.

	\textbf{Bayesian Network.}
	In general, a Probabilistic Graphical Model~\cite{koller2009probabilistic}, and specifically a \textit{Bayesian network}~\cite{pearl1985bayesian,chavira2008probabilistic}, encodes the dependencies and conditional independence between a set of random variables. In this paper, we leverage an access to a Bayesian network on $ \nonsensitive \cup \sensitive $ that represents the joint distribution on them. 	A Bayesian network is denoted by a pair $ (\graph, \theta)$, where $ \graph \triangleq (\mathbf{V}, \mathbf{E}) $ is a DAG (Directed Acyclic Graph), and $\theta$ is a set of parameters encoding the conditional probabilities induced by the joint distribution under investigation. Each vertex $V_i \in \mathbf{V}$ corresponds to a random variable. Edges $ \mathbf{E} \in \mathbf{V} \times \mathbf{V} $ imply conditional dependencies among variables. For each variable $ V_i \in \mathbf{V} $, let $ \parent(V_i) \subseteq \mathbf{V} \setminus \{V_i\} $ denote the set of parents of $ V_i $. Given $\parent(V_i)$ and parameters $\theta$, $ V_i $ is independent of its other non-descendant variables in $\graph$. Thus, for the assignment $ v_i $ of $ V_i $ and $ \mathbf{u} $ of $ \parent(V_i) $, the aforementioned semantics of a Bayesian network encodes the joint distribution of $\mathbf{V}$ as:

	\begin{equation}
		\begin{split}
				\Pr[V_1=v_1, \dots, &V_{|\mathbf{V}|}=v_{|\mathbf{V}|}] = \\ 
			&\prod_{i=1}^{|\mathbf{V}|} \Pr[V_i = v_i | \parent(V_i) = \mathbf{u}; \theta].
		\end{split}
	\label{fvgm_eq:BN}
	\end{equation}

### 2022-04-07 chapters/fairness/fvgm/conclusion.tex
updated files

\section{Conclusion}
In this paper, we propose {\fvgm}, an efficient fairness verification framework for linear classifiers based on a novel stochastic subset-sum problem. {\fvgm} encodes a graphical model of feature-correlations, represented as a Bayesian Network, and computes multiple group and causal fairness metrics accurately. We experimentally demonstrate that {\fvgm} is not only more accurate and scalable than the existing verifiers but also applicable in practical fairness tasks, such as verifying fairness attacks and enhancing algorithms, and computing the fairness influence functions. 
As a future work, we aim to design fairness-enhancing algorithms certified by fairness verifiers, such as {\fvgm}. %We also investigate a principled approach to detect the sources of bias by computing fairness influence functions (FIF). 
Since {\fvgm} serves as an accurate and scalable fairness verifier for linear classifiers, it will be interesting to design such verifiers for other ML models.

>>>>

\section{Conclusion}
In this paper, we propose {\fvgm}, an efficient fairness verification framework for linear classifiers based on a novel stochastic subset-sum problem. {\fvgm} encodes a graphical model of feature-correlations, represented as a Bayesian Network, and computes multiple group and causal fairness metrics accurately. We experimentally demonstrate that {\fvgm} is not only more accurate and scalable than the existing verifiers but also applicable in practical fairness tasks, such as verifying fairness attacks and enhancing algorithms, and computing the fairness influence functions. 
As a future work, we aim to design fairness-enhancing algorithms certified by fairness verifiers, such as {\fvgm}. %We also investigate a principled approach to detect the sources of bias by computing fairness influence functions (FIF). 
Since {\fvgm} serves as an accurate and scalable fairness verifier for linear classifiers, it will be interesting to design such verifiers for other ML models.
%Though the proposed solution with a Bayesian network has exponential worst-case complexity, it would be interesting to develop a pseudo-polynomial approximation in the future.

\iffalse
\section*{Acknowledgments}
This work was supported in part by National Research Foundation, Singapore under its NRF Fellowship Programme [NRF-NRFFAI1-2019-0004 ] and AI Singapore Programme [AISG-RP-2018-005],  and NUS ODPRT Grant [R-252-000-685-13]. The computational work for this paper was performed on resources of Max Planck Institute for Software Systems, Germany and the National Supercomputing Centre, Singapore (\url{https://www.nscc.sg}).
\fi
